---
title: "Restoration Ecology Pub Nyen"
author: "GN"
date: "2025-06-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#------------------------------------------------------------------------------#
#                          Jones Center at Ichauway                            #
#                                                                              #
# Title:           Gabe Nyen data analysis                                     #
# Authors:         Gabe Nyen, Lisa Giencke                                     #
# Last updated:    September 17, 2024                                          #
# Last updated by: lmg                                                         #
# Objective:       Examine differences in leave and take out rows with and     # 
#                  without seeded ground cover species                         #
#                  following Hurricane Micheal                                 #
#                                                                              #
#------------------------------------------------------------------------------#

library(tidyverse)
library(vegan)
library(broom)
library(janitor)
library(ggplot2)
library(ggrepel)
library(indicspecies)
library(glmmADMB)
library(broom)
library(vegan)
library(parallel)
library(janitor)
library("latex2exp")
library("foreign")
library(MuMIn)
SEED <- 1900
set.seed(SEED)

rm(list=ls(all=T))
```


```{r}
#setwd("R:/plant_ecology/misc_project_directories/nyen")
#metadata<- read.csv("C:/Users/gaben/Downloads/gn_metadata_wzeros.csv", header=T, na.strings="")
metadata<- read.csv("C:/Users/nyen0010/OneDrive - Joseph W. Jones Ecological Research Center/gn_metadata_wzeros.csv", header=T, na.strings="")
#------------------------Choose analysis to run---------------------------------

#Choose input dataset (dat1 = all plot weights; dat2 = plot weights combined by 
#stand/row type)
input_dat_choice <- 1

#----------------------Read dataset and dataset prep----------------------------

#You WILL HAVE TO CANGE THIS TO YOUR COMUPTER LOCATION
#metadata<- read.csv("gn_metadata_wzeros.csv", header=T, na.strings="")
#import the metadata, make all empty entries into NA's

#Check out data formatting - lmg
str(metadata)
summary(metadata)

#This is listed as Aristida fasciculata, but that's not a species.
#I made the assumption that it was meant to be Agalinis fasciculata. - lmg
metadata$species_code[which(metadata$species_code == "ARFA")] <- "AGFA"

#Both codes are used in dataset, but are for the same species. - lmg
metadata$species_code[which(metadata$species_code == "ERST")] <- "ERSG"

#Both codes are used in dataset, but are for the same species. - lmg
metadata$species_code[which(metadata$species_code == "LACA")] <- "LACD"

#Listed as Rhnychosia deformus in the dataset, but I assume it to be 
#Rhynchosia difformis - lmg
metadata$species_code[which(metadata$species_code == "RHDE")] <- "RHDI"

# I think I remember looking at this with Alisen in 2023, but RHMI (even though
# it's on the Ichauway species list), is not considered to occur in this part of 
# Georgia). I assume it to be RHDI. - lmg

metadata$species_code[which(metadata$species_code == "RHMI")] <- "RHDI"
  
metadata <- metadata %>% mutate(species_code = coalesce(species_code, species))
head(metadata, 20)
#replace all blank species with the unknown official names listed in the species_code column
```


```{r}
#Make sure all listed species codes look ok - lmg
unique(metadata$species_code)[order(unique(metadata$species_code))]

data <-metadata %>% filter(!is.na(species_code))
#get rid of the entries without a species code after the coalesce.

data <- data %>% filter(!plot_id=="9-6-K")
#takes out a plot that didnt have data to it
```

## Keep stand 3 in the dataset
The following chunk will not be run - GN
```{r}
#Remove stand 3 because it's so compositionally different - lmg
#data <- data %>% filter(!stand==3)
```


```{r}
data<- data  %>% replace_na(list(row = "L"))

#Log transform weight data - lmg
data$wt_trans <- log10(data$dried_weight_g+0.01)+2

#Look at data distributions - lmg
#ggplot(data, aes(x=dried_weight_g)) + geom_histogram(bins = 25)
#ggplot(data, aes(x=wt_trans)) + geom_histogram(bins = 25)


varibs <- data %>% summarise(plot_id, n_seedlings, openness, lai_2000_lin, 
                             ppfd_total_under_per_day_m_jor_mol_m2day, 
                             gap_fraction, row, stand, transect) %>%
  distinct()
#create a variable data frame
#I dont think that the error code maters


varibs<- varibs  %>% replace_na(list(n_seedlings=0, openness=0, lai_2000_lin = 0, 
                                     ppfd_total_under_per_day_m_jor_mol_m2day = 0, 
                                     gap_fraction= 0, row = "L", stand=0, standbyrow= 0))
#changes the NA values to be the resting values of zero or in the row it changes to a Leave row
#removed plot_id to avoid error message - lmg

#Visually plot environmental variables against one another - lmg
pairs(varibs[,2:6])
```


```{r}
#Create dataset of weights for each stand/row type combination
data.stand.row <- aggregate(data$dried_weight_g, by=list(data$row, data$stand, 
                                                         data$species_code), sum)
names(data.stand.row) <- c("row", "stand", "species_code", "dried_weight_g")

#Calculate the number of plots per stand/row type combination - lmg
plots <- unique(data[c("stand", "plot_id", "row")])
plot.counts <- aggregate(plots$stand, by=list(plots$row, plots$stand), length)
names(plot.counts) <- c("row", "stand", "num_plots")

#Combine weight and plot number datasets to get a mean weight per stand/
#row type combination, then transform mean weight data
data.stand.row <- merge(data.stand.row, plot.counts, by = c("stand", "row"))
data.stand.row$mean.wt <- data.stand.row$dried_weight_g/data.stand.row$num_plots
data.stand.row <- data.stand.row[order(data.stand.row$stand, data.stand.row$row, 
                                       data.stand.row$species_code, decreasing = F), ]
data.stand.row$wt_trans <- log10(data.stand.row$dried_weight_g+0.0001)+3

#Look at data distributions - lmg
ggplot(data.stand.row, aes(x=dried_weight_g)) + geom_histogram(bins = 25)
ggplot(data.stand.row, aes(x=wt_trans)) + geom_histogram(bins = 25)
```
# the following pivot table uses wt_trans as apposed to non transformed data.
Must follow up to test if the wt_trans is 1.) needed, 2.) still significant
3.) if all other tests use the same wt_trans data.


```{r}
dat1 <- data %>% dplyr::select(plot_id, species_code, wt_trans) %>%
  pivot_wider(names_from = species_code, values_from = wt_trans, values_fn = sum, values_fill = 0)
#this transforms the data frame into a wide format such that it can be used in NMDS
#view(dat1) # dat1 is now ready to go into an NMDS engine 

dat2 <- data.stand.row %>% dplyr::select(stand, row, species_code, wt_trans) %>%
  pivot_wider(names_from = species_code, values_from = wt_trans, values_fn = sum, values_fill = 0)
#this transforms the data frame into a wide format such that it can be used in NMDS - lmg
```
Here, dat2 is just stands and row types to feed into an NMNDS

```{r}
#Create dataset of environmental variables to pair with stand/row type weight 
#data - lmg
varibs.stand.row <- aggregate(varibs[,c(2:6)], by=list(varibs$row, varibs$stand), mean)
names(varibs.stand.row)[1:2] <- c("row", "stand")
varibs.stand.row$total.biomass <- aggregate(data.stand.row$mean.wt, 
                                            by=list(data.stand.row$row, 
                                                    data.stand.row$stand), sum)[,3]
varibs.stand.row$total.biomass.trans <- log10(varibs.stand.row$total.biomass)
```
# Unsure if the above chunk is needed.. 

```{r}
#Create dataset of seeded weights - lmg
seeded.weight <- aggregate(subset(data.stand.row$mean.wt, 
                                  data.stand.row$species_code %in% 
                                    c("ARBE", "PAVI", "SONU", "SOSE", "TEVI")),
                           by=list(subset(data.stand.row$row, data.stand.row$species_code %in% 
                                            c("ARBE", "PAVI", "SONU", "SOSE", "TEVI")),
                                   subset(data.stand.row$stand, data.stand.row$species_code %in% 
                                            c("ARBE", "PAVI", "SONU", "SOSE", "TEVI"))), sum)
names(seeded.weight) <- c("row", "stand", "seeded.biomass")

#Combine environmental variables and seeded weights, then transform seeded
#weights, and add treatment (seeded/unseeded) - lmg
varibs.stand.row <- merge(varibs.stand.row, seeded.weight, by=c("stand", "row"), all.x = T)
varibs.stand.row<- varibs.stand.row  %>% replace_na(list(seeded.biomass=0))
varibs.stand.row$seeded.biomass.trans <- log10(varibs.stand.row$seeded.biomass+0.01)+2
varibs.stand.row$trt <- ifelse(varibs.stand.row$stand %in% 6:9, "Seeded", "Unseeded")

#Once again visualize data - lmg
#ggplot(varibs.stand.row, aes(x=sqrt(total.biomass.trans))) + geom_histogram(bins = 25)

#Visualizing/testing space for environmental variables
#boxplot(varibs.stand.row$seeded.biomass ~ interaction(varibs.stand.row$row, varibs.stand.row$trt))  
#summary(aov(varibs.stand.row$seeded.biomass ~ as.factor(varibs.stand.row$row)*as.factor(varibs.stand.row$trt)))))
#TukeyHSD(aov(varibs.stand.row$seeded.biomass ~ as.factor(varibs.stand.row$row)*as.factor(varibs.stand.row$trt)))))



if(input_dat_choice ==1){
  input_dat <- dat1 %>% column_to_rownames("plot_id")
}
if(input_dat_choice == 2){
  input_dat <- as.data.frame(dat2[,3:dim(dat2)[2]])
}

#Create presence dataset to be used in rare species removals
input_dat.pa <- input_dat > 0
input_dat.pa[which(input_dat.pa == "TRUE")] <- 1

#Remove rare species (those that occur in less than 3 stand/row type combinations
#from dataset
input_dat.no.rare <- input_dat
input_dat.no.rare[dim(input_dat)[1]+1,] <- colSums(input_dat.pa[,] > 0)
input_dat.no.rare <- input_dat.no.rare[1:(dim(input_dat)[1]), input_dat.no.rare[dim(input_dat)[1] + 1,] > 2]

rm(input_dat.pa, input_dat)

#--------------------------------Run ordination---------------------------------
nmdsE <- metaMDS(comm = input_dat.no.rare,
                 autotransform = FALSE,
                 distance = "bray",
                 engine = "monoMDS",
                 k = 3,
                 #parallel = clus, #this is only needed when using 
                 #all of the computer cores acting as clusters. 
                 #Need a ton of cores before this is useful!
                 weakties = TRUE,
                 model = "global",
                 maxit = 5000,
                 try = 2000, # Use 2000, 200 for quick process
                 trymax = 2000) #Use 2000, 200 for quick process
nmdsE
nmdsE$stress


#----------------Create dataset for plotting ordination results-----------------
#Choose which plot identifiers to pull based on input dataset choice - lmg
if(input_dat_choice == 1){
  plot.dat <- as.data.frame(cbind(varibs$stand, varibs$row, (nmdsE$points)))
}
if(input_dat_choice == 2){
  plot.dat <- as.data.frame(cbind(varibs.stand.row$stand, varibs.stand.row$row, 
                                (nmdsE$points)))
}

names(plot.dat)[1:2] <- c("Stand", "RowType")
plot.dat[,3] <- as.numeric(plot.dat[,3])
plot.dat[,4] <- as.numeric(plot.dat[,4])
plot.dat[,5] <- as.numeric(plot.dat[,5])
plot.dat$Trt <- ifelse(plot.dat$Stand %in% 6:9, "Seeded", "Unseeded")
plot.dat$Trt.Row <- interaction(plot.dat$Trt, plot.dat$RowType)
plot.dat$Trt.Row <- factor(plot.dat$Trt.Row, levels = c("Seeded.L", "Seeded.TO",
                                                        "Unseeded.L", "Unseeded.TO"))
```


```{r}
#--------------Extract significant species vectors for plotting-----------------

#Calculate species vectors - lmg
spp.vect <- envfit(nmdsE, input_dat.no.rare, permutations = 999, choices = c(1:2))

vec.ggplot <- as.data.frame(spp.vect$vectors$arrows*sqrt(spp.vect$vectors$r))
vec.ggplot$species <- rownames(vec.ggplot)

#Keep only significant species - lmg
sig.species <- as.list(spp.vect$vectors)

#creating dataset of species vectors and p values - lmg
pvals <- as.data.frame(sig.species$pvals)
arrows <- as.data.frame(sig.species$arrows*sqrt(sig.species$r))
spp.ord <- cbind(arrows, pvals)

#Subset to just significant species - lmg
spp.ord.sig <-subset(spp.ord,pvals<0.05)
spp.ord.sig <- cbind(spp.ord.sig, Species = rownames(spp.ord.sig))

rm(spp.vect, vec.ggplot, sig.species, pvals, arrows, spp.ord)

#-------------------
# extract the seeded species to put them into different color in the ordination
spp.ord.sig.seed <- filter(spp.ord.sig, Species == c("TEVI"))
spp.ord.sig.seed1 <- filter(spp.ord.sig, Species == c("SONU"))
spp.ord.sig.seed2 <- filter(spp.ord.sig, Species == c("SOSE"))


```
 # spp.ord.sig now contains all spp vectors that are significant to the ordination



Line right below this is where Gap fraction should be placed along with c(2,4,"6")
```{r}
#--------Extract significant environmental vectors for plotting-----------------



#Choose environmental variables depending on choice of input dataset - lmg
if(input_dat_choice == 1){
  names(varibs)[c(2,4)] <- c("Number of Seedlings", "LAI 2000", "Gap Fraction")
  env.vect <- envfit(nmdsE, varibs[,c(2,4,6)], permutations = 999, na.rm = T)
}
if(input_dat_choice == 2){
  names(varibs.stand.row)[c(3,4,9,11)] <- c("Number of Seedlings", "Openness", "Total Biomass", "Seeded Biomass")
  env.vect <- envfit(nmdsE, varibs.stand.row[,c(3,4,9,11)], permutations = 999, na.rm = T)
}

vec.ggplot<-as.data.frame(env.vect$vectors$arrows*sqrt(env.vect$vectors$r))
vec.ggplot$species<-rownames(vec.ggplot)

#Keep only significant environmental variables - lmg
sig.env <- as.list(env.vect$vectors)

#create dataset of environmental variable vectors and p values - lmg
pvals <- as.data.frame(sig.env$pvals)
arrows <- as.data.frame(sig.env$arrows*sqrt(sig.env$r))
env.ord <- cbind(arrows, pvals)

#Subset to just significant environmental variables - lmg
env.ord.sig<-subset(env.ord,pvals<0.05)
env.ord.sig <- cbind(env.ord.sig, Variables = rownames(env.ord.sig))

rm(env.vect, vec.ggplot, sig.env, pvals, arrows, env.ord)
```


```{r}
#----------Create function to calculate confidence ellipses---------------------


#Prettied up - species
# function for ellipsess - just run this, is used later
#taken from the excellent stackoverflow Q+A: http://stackoverflow.com/questions/13794419/plotting-ordiellipse-function-from-vegan-package-onto-nmds-plot-created-in-ggplo
veganCovEllipse <- function (cov, center = c(0, 0), scale = 1, npoints = 100) 
{
  theta <- (0:npoints) * 2 * pi/npoints
  Circle <- cbind(cos(theta), sin(theta))
  t(center + scale * t(Circle %*% chol(cov)))
}

df.ell <- data.frame()
for(i in unique(plot.dat$Trt.Row)){
  df.ell <- rbind(df.ell, cbind(as.data.frame(with(plot.dat[plot.dat$Trt.Row == i,],
                                                   veganCovEllipse(cov.wt(cbind(MDS1,MDS2),wt=rep(1/length(MDS1),length(MDS1)))$cov,center=c(mean(MDS1),mean(MDS2))))),
                                Clusters=i))
}


df.ell$Clusters <- factor(df.ell$Clusters, levels = c("Seeded.L", "Seeded.TO",
                                                       "Unseeded.L", "Unseeded.TO"))
```


```{r}
#-----------------------------Plot ordination-----------------------------------

ggplot(data = plot.dat, aes(x=MDS1, y=MDS2)) + theme_bw() + 
  geom_point(data = plot.dat, aes(fill = Trt.Row), size = 4, pch = 22) + 
  geom_segment(data=spp.ord.sig,aes(x=0,xend=NMDS1,y=0,yend=NMDS2),
               arrow = arrow(length = unit(0.3, "cm")),colour="grey64", size = 0.8,
              inherit.aes=FALSE) + 
  geom_segment(data=env.ord.sig,aes(x=0,xend=NMDS1,y=0,yend=NMDS2),
               arrow = arrow(length = unit(0.3, "cm")),colour="black", size = 0.8,
               inherit.aes=FALSE) +   
  geom_path(data = df.ell, aes(x=MDS1, y=MDS2, color = as.factor(Clusters), 
                               group = as.factor(Clusters)), inherit.aes = F, 
            size = 1.5) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  geom_text_repel(data=spp.ord.sig,aes(x=NMDS1,y=NMDS2,label=Species),size=4, color = "black", 
                  inherit.aes = F, hjust = ifelse(spp.ord.sig$NMDS1 < 0, 1.2, 1)) +
  geom_text_repel(data=env.ord.sig,aes(x=NMDS1,y=NMDS2,label=Variables),size=4, color = "black", 
                  inherit.aes = F, hjust = ifelse(env.ord.sig$NMDS1 < 0, 1.2, 1)) +
  scale_color_brewer(palette = "Spectral", name = "95% Confidence Ellipse", 
                     labels = c("Seeded + Leave", "Seeded + Take Out", 
                                "Unseeded + Leave", "Unseeded + Take Out")) +
  scale_fill_brewer(palette = "Spectral", name = "Plot Type",
                    labels = c("Seeded + Leave", "Seeded + Take Out", 
                               "Unseeded + Leave", "Unseeded + Take Out")) +
  guides(fill = guide_legend(order = 1), col = guide_legend(order = 2)) +
  theme(legend.position = "right") 

#------------------------Second Plot view---------------------------------------

ggplot(data = plot.dat, aes(x=MDS1, y=MDS2)) + theme_bw() + 
  geom_point(data = plot.dat, aes(fill = Trt.Row), alpha=0.65, size = 4, pch = 21) + 
  geom_segment(data=spp.ord.sig,aes(x=0,xend=NMDS1,y=0,yend=NMDS2),
               arrow = arrow(length = unit(0.3, "cm")),colour="grey64", size = 0.8,
               inherit.aes=FALSE) + 
  geom_segment(data=spp.ord.sig.seed,aes(x=0,xend=NMDS1,y=0,yend=NMDS2), #seeded spp
               arrow = arrow(length = unit(0.3, "cm")),colour="red4", size = 1.1,
              inherit.aes=FALSE) +
  geom_segment(data=spp.ord.sig.seed1,aes(x=0,xend=NMDS1,y=0,yend=NMDS2), #seeded spp
               arrow = arrow(length = unit(0.3, "cm")),colour="red4", size = 1.1,
              inherit.aes=FALSE) +
  geom_segment(data=spp.ord.sig.seed2,aes(x=0,xend=NMDS1,y=0,yend=NMDS2), #seeded spp
               arrow = arrow(length = unit(0.3, "cm")),colour="red4", size = 1.1,
              inherit.aes=FALSE) +
  geom_segment(data=env.ord.sig,aes(x=0,xend=NMDS1,y=0,yend=NMDS2),
               arrow = arrow(length = unit(0.3, "cm")),colour="black", size = 0.8,
               inherit.aes=FALSE) +   
  #GN # geom_path(data = df.ell, aes(x=MDS1, y=MDS2, color = as.factor(Clusters), 
   #GN#                            group = as.factor(Clusters)), inherit.aes = F, 
   #GN#         size = 1.5) + 
  #GN?#      theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + #GN hashtaged this out to allow it to render. Unknown issue
  geom_text_repel(data=spp.ord.sig,aes(x=NMDS1,y=NMDS2,label=Species),size=4, color = "black", 
                  inherit.aes = F, hjust = ifelse(spp.ord.sig$NMDS1 < 0, 1.2, 1)) +
  geom_text_repel(data=env.ord.sig,aes(x=NMDS1,y=NMDS2,label=Variables),size=5, color = "black", 
                  inherit.aes = F, hjust = ifelse(env.ord.sig$NMDS1 < 0, 1.2, 1)) +
  scale_color_brewer(palette = "Spectral", name = "95% Confidence Ellipse", 
                     labels = c("Seeded + Leave", "Seeded + Take Out", 
                                "Unseeded + Leave", "Unseeded + Take Out")) +
  scale_fill_brewer(palette = "Spectral", name = "Plot Type",
                    labels = c("Seeded + Leave", "Seeded + Take Out", 
                               "Unseeded + Leave", "Unseeded + Take Out")) +
  guides(fill = guide_legend(order = 1), col = guide_legend(order = 2)) +
  theme(legend.position = "right") 



# Gabe - Experiment with only keeping the lables for the sig species on the plot and the envionmental vectors. That may look cleaner





#------------------------Second Plot view---------------------------------------
```


```{r}
#------------------------Indicator species analysis-----------------------------


ind.spp <- multipatt(input_dat.no.rare, plot.dat$Trt.Row)
summary(ind.spp)


###====================Gabe Tests==========================
#prep a dried weight vector# 
dat1_long <- dat1 %>%         #take the one almost ready for nmds and long it
  pivot_longer(!plot_id, names_to = "species", values_to = "biomass")

dat1_summary <- dat1_long %>% #take that and cluster add it to plot
  group_by(plot_id) %>%
  summarise(biomass_sum = sum(biomass))
#add up spp weights to pool for each plot

dat1_long1 <- dat1_long %>% 
  mutate(species = as.character(species))


dat1_long_noPIPA <- dat1_long1 %>% filter(species != "PIPA")

dat1_summary_noPIPA <- dat1_long_noPIPA %>% 
  group_by(plot_id) %>%
  summarise(biomass_sum = sum(biomass))

#this is a df with species weight on the inside and plot & species as headers
input_dat <- dat1 %>% column_to_rownames("plot_id")

#prep a dried weight vector#

adonis2(input_dat.no.rare ~ Trt,
        plot.dat,
        permutations = 99999,
        method = "bray",
        sqrt.dist = F,
        by = "terms")

view(plot.dat)

varibs_test <- varibs
column_to_rownames(varibs_test, var = "plot_id")
varibs_test <- varibs_test %>% clean_names()


ef_test<- envfit(nmdsE ~ Trt, plot.dat, permutations = 99999, choices = 1:3, na.rm = T, iterations = 1000)
ef_test

ef_test1<- envfit(nmdsE ~ openness, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations = 1000)
ef_test1

ef <- envfit(nmdsE ~ lai_2000_lin, varibs_test, permutations = 99999, choices = 1:3, na.rm=T, iterations = 1000)
ef

ef1 <- envfit(nmdsE ~ stand, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations= 1000)
ef1 #Significant ! ! !

ef2 <- envfit(nmdsE ~ n_seedlings, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations= 1000)
ef2

ef3 <- envfit(nmdsE ~ seed_trt, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations= 1000)
ef3 #Significant ! ! !
# how was this tested? what does this test do? look at : https://www.flutterbys.com.au/stats/tut/tut15.1.html

efstand <- envfit(nmdsE ~ stand, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations= 1000)
efstand

ef4 <- envfit(nmdsE ~ ppfd_total_under_per_day_m_jor_mol_m2day, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations= 1000)
ef4

ef5 <- envfit(nmdsE ~ openness, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations= 1000)
ef5

ef6 <- envfit(nmdsE ~ gap_fraction, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations= 1000)
ef6

ef7 <- envfit(nmdsE ~ biomass_sum, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations= 1000)
ef7

ef8 <- envfit(nmdsE ~ row, varibs_test, permutations = 99999, choices = 1:3, na.rm = T, iterations= 1000)
ef8

f <-tibble(ef$vectors$arrows[1:3]) %>% rename(lai_2000_lin= "ef$vectors$arrows[1:3]")
f1<-tibble(ef1$vectors$arrows[1:3]) %>% rename(stand= "ef1$vectors$arrows[1:3]")
f2<-tibble(ef2$vectors$arrows[1:3]) %>% rename(n_seedlings= "ef2$vectors$arrows[1:3]")
#f3<-tibble(ef3$vectors$arrows[1:3]) %>% rename(seed_trt= "ef3$vectors$arrows[1:3]")
f4<-tibble(ef4$vectors$arrows[1:3]) %>% 
  rename(ppfd_total_under_per_day_m_jor_mol_m2day= "ef4$vectors$arrows[1:3]")
f5<-tibble(ef5$vectors$arrows[1:3]) %>% rename(openness= "ef5$vectors$arrows[1:3]")
f6<-tibble(ef6$vectors$arrows[1:3]) %>% rename(gap_fraction= "ef6$vectors$arrows[1:3]")
f7<-tibble(ef7$vectors$arrows[1:3]) %>% rename(biomass_sum= "ef7$vectors$arrows[1:3]")
#f8<-tibble(ef8[["factors"]][["centroids"]]) %>% rename(row= ef8[["factors"]][["centroids"]])


flist <- bind_cols(f,f1,f2,f4,f5,f6,f7) %>% as.data.frame()
rownames(flist) = c("NMDS1", 
                    "NMDS2",
                    "NMDS3")
flist

rank_flist<-abs(flist)
max(rank_flist[1,]) # Stand is most highly corelated with NMDS1
max(rank_flist[2,]) # n_seedlings is most highly corelated with NMDS2
max(rank_flist[3,]) # biomass_sum is most highly corelated with NMDS3


#write.csv(flist, "C:/Users/gn238/Downloads/vector_associations.csv", row.names=T)


#--------------------------------The end----------------------------------------
```

